Figures from: Vertebrate host availability and the thermal properties of mosquito-borne parasite transmission

### 2023-05-22 manuscript-figures.Rmd
Update manuscript-figures.Rmd

# Define temperature range of study
Temps <- seq(10, 40, length.out = 601) # full: length.out = 601, thin: length.out = 301

# Thin sample size
thin_size <- 1000 # full = 1000, thin = 100

plot_bool = FALSE
source("code/trait-transform.R", local = TRUE)
# Host density vector: Number of values to include to consider for vertebrate host density
KH_vec_length <- 300 # full = 300, thin = 20

# Biting tolerance vector: Number of values to consider for biting tolerance
sigmaH_vec_length <- 300 # full = 300, thin = 20

# Get data for plotting thermal performance curves (data.Vec)
source("code/trait-variation.R", local = TRUE)

@@

# R0 data
data.R0 <- read_rds("results/R0_vals.rds") # !!! needs to be updated to include 1e-1, 1e0, 1e1
# Topt data
data.Topt <- read_rds("results/Topt_vals.rds")
# Critical thermal extrema data
data.CT <- read_rds("results/CT_vals.rds")

@@

data.R0_norm <- data.R0 %>% 
  filter(KH %in% newKH_vals) %>% 
  filter(variable == "R0") %>%
  select(-variable) %>%
  # normalize variables to have a maximum of 1
  group_by(system_ID, Model, sigmaH, KH) %>% 
  # mutate(mean_norm = mean / max(mean)) %>% 
  mutate(median_norm = median / max(median)) %>%
  # mutate(highHCI_norm = highHCI / max(median)) %>% 
  # mutate(lowHCI_norm = lowHCI / max(median)) %>% 
  arrange(system_ID, sigmaH, KH, Temperature, median_norm)

R0_plot <- ggplot(mapping = aes(x = Temperature, group = KH)) +
  # path of mean, normalized R0 as a function of temperature (finite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == 100),
    aes(y = median_norm, colour = KH, linetype = "finite"),
    lwd = 1
  ) +
  # path of mean, normalized R0 as a function of temperature (infinite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == Inf),
    aes(y = median_norm, linetype = "infinite", colour = "black"),
    colour = "black",
    lwd = 1.5
  ) +
  # # 89% HCI of R0 TPC curves
  # geom_ribbon(
  #   data = filter(data.R0_norm, sigmaH == 100),
  #   aes(ymin = lowHCI_norm, ymax = highHCI_norm, fill = KH),
  #   alpha = 0.1
  # ) +
  # x-axis:
  scale_x_continuous(
    name = "Temperature (C)",
    expand = c(0, 0),
    limits = c(14, 34)
  ) +
  # y-axis:
  scale_y_continuous(
    name = TeX("Normalized $R_0$"),
    expand = c(0, .05)
  ) +
  # color: scaled log10, color-blind friendly
  scale_color_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # fill: scaled log10, color-blind friendly
  scale_fill_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # linetype:
  scale_linetype_manual(
    name = "Biting tolerance\n(bites per host per day)",
    breaks = c("finite", "infinite"),
    labels = unname(c(TeX("100 (Chitnis)"), TeX("$\\infty$ (Ross-Macdonald)"))),
    values = c("solid", "22")
  ) +
  # faceting:
  facet_wrap(~system_ID, scales = "free", nrow = 2, labeller = labeller()) +
  # legend
  guides(
    color = guide_colourbar(
      title.position = "top",
      nrow = 1,
      barwidth = 12,
      show.limits = TRUE,
      direction = "horizontal"
    ),
    linetype = guide_legend(
      title.position = "top",
      nrow = 2,
      keywidth = 4,
      direction = "vertical"
    )
  ) +
  # theme options:
  theme_minimal_vgrid(16) +
  theme(
    legend.box = "vertical",
    legend.margin = margin(),
    legend.title = element_text(size = 10),
    legend.text = element_text(
      size = 10,
      hjust = 0
    ),
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

@@

Toptalt_df <- data.Topt %>%
  ungroup() %>%
  distinct() %>%
  # Arrange along the plotting variables
  arrange(system_ID, sigmaH, KH, mean) %>% 
  mutate(sigmaH_proxy = sigmaH)

@@

  # fill:
  scale_fill_manual(
    name = "Transmission\nthermal optimum (°C)",
    values = met.brewer("Johnson",
                        n = length(Topt_breaks),
                        direction = -1
    )
  ) +
  # color:
  scale_colour_manual(
    name = "",
    labels = c("R0=1" = unname(TeX("$R_0=1$"))),
    values = c("black"),
    guide = guide_legend(
      order = 1,
      keyheight = unit(1, "cm"),
      override.aes = list(size = 6)
    )
  ) +
  # faceting:
  facet_wrap(~system_ID, scales = "free", nrow = 2, labeller = labeller()) +
  # legend:
  guides(fill = guide_coloursteps(
    title.position = "top", title.hjust = 0,
    barheight = unit(5, "cm"),
    show.limits = TRUE
  )) +
  # theme options:
  theme_minimal(16) +
  theme(
    legend.box = "vertical",
    legend.position = "right",
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.direction = "vertical",
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

>>>>

# # Define temperature range of study
# Temps <- seq(10, 40, length.out = 601) # full: length.out = 601, thin: length.out = 301
# 
# # Thin sample size
# thin_size <- 1000 # full = 1000, thin = 100
# 
# plot_bool = FALSE
# source("code/trait-transform.R", local = TRUE)
# # Host density vector: Number of values to include to consider for vertebrate host density
# KH_vec_length <- 300 # full = 300, thin = 20
# 
# # Biting tolerance vector: Number of values to consider for biting tolerance
# sigmaH_vec_length <- 300 # full = 300, thin = 20
# 
# # Get data for plotting thermal performance curves (data.Vec)
# !!! save a specific dataset to use for plotting trait TPCs
# source("code/trait-variation.R", local = TRUE)

@@

# R0 data # !!! needs to be updated to include 1e-1, 1e0, 1e1
data.R0 <- read_rds("results/R0_vals.rds")
# Topt data
data.Topt <- read_rds("results/Topt_vals.rds")
# Critical thermal extrema data
data.CT <- read_rds("results/CT_vals.rds")

@@

data.R0_norm <- data.R0 %>% 
  ungroup() %>% 
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* & DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* & ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c) *Ae. albopictus* & DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *Anopheles* spp. & *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* & WNV"
  )) %>% 
  filter(KH %in% newKH_vals) %>% 
  filter(variable == "R0") %>%
  select(-variable) %>%
  # normalize variables to have a maximum of 1
  group_by(system_ID, Model, sigmaH, KH) %>% 
  # mutate(mean_norm = mean / max(mean)) %>% 
  mutate(median_norm = median / max(median)) %>%
  # mutate(highHCI_norm = highHCI / max(median)) %>% 
  # mutate(lowHCI_norm = lowHCI / max(median)) %>% 
  arrange(system_ID, sigmaH, KH, Temperature, median_norm)

R0_plot <- ggplot(mapping = aes(x = Temperature, group = KH)) +
  # path of mean, normalized R0 as a function of temperature (finite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == 100),
    aes(y = median_norm, colour = KH, linetype = "finite"),
    lwd = 1
  ) +
  # path of mean, normalized R0 as a function of temperature (infinite sigmaH):
  geom_path(
    data = filter(data.R0_norm, sigmaH == Inf),
    aes(y = median_norm, linetype = "infinite", colour = "black"),
    colour = "black",
    lwd = 1.5
  ) +
  # # 89% HCI of R0 TPC curves
  # geom_ribbon(
  #   data = filter(data.R0_norm, sigmaH == 100),
  #   aes(ymin = lowHCI_norm, ymax = highHCI_norm, fill = KH),
  #   alpha = 0.1
  # ) +
  # x-axis:
  scale_x_continuous(
    name = "Temperature (°C)",
    expand = c(0, 0),
    limits = c(14, 34)
  ) +
  # y-axis:
  scale_y_continuous(
    name = TeX("Normalized $R_0$"),
    expand = c(0, .05)
  ) +
  # color: scaled log10, color-blind friendly
  scale_color_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # fill: scaled log10, color-blind friendly
  scale_fill_viridis(
    name = "Vertebrate host population\ndensity (ind/ha)",
    trans = "log10",
    breaks = 10^seq(-2, 4),
    labels = unname(c(0.01, 0.1, 1, 10, 100, TeX("$10^3$"), TeX("$10^4$"))),
    option = "plasma"
  ) +
  # linetype:
  scale_linetype_manual(
    name = "Biting tolerance\n(bites per host per day)",
    breaks = c("finite", "infinite"),
    labels = unname(c(TeX("100 (Chitnis)"), TeX("$\\infty$ (Ross-Macdonald)"))),
    values = c("solid", "22")
  ) +
  # faceting:
  facet_wrap(~system_label, scales = "free", nrow = 2, labeller = labeller()) +
  # legend
  guides(
    color = guide_colourbar(
      title.position = "top",
      nrow = 1,
      barwidth = 12,
      show.limits = TRUE,
      direction = "horizontal"
    ),
    linetype = guide_legend(
      title.position = "top",
      nrow = 2,
      keywidth = 4,
      direction = "vertical"
    )
  ) +
  # theme options:
  theme_minimal_vgrid(16) +
  theme(
    legend.box = "vertical",
    legend.margin = margin(),
    legend.title = element_text(size = 10),
    legend.text = element_text(
      size = 10,
      hjust = 0
    ),
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

@@

Toptalt_df <- data.Topt %>%
  ungroup() %>%
  mutate(system_label = case_when(
    system_ID == "Aedes aegypti / DENV" ~ "(a) *Ae. aegypti* & DENV",
    system_ID == "Aedes aegypti / ZIKV" ~ "(b) *Ae. aegypti* & ZIKV",
    system_ID == "Aedes albopictus / DENV" ~ "(c) *Ae. albopictus* & DENV",
    system_ID == "Anopheles gambiae / Plasmodium falciparum" ~ "(d) *Anopheles* spp. & *P. falciparum*",
    system_ID == "Culex quinquefasciatus / WNV" ~ "(e) *Cx. quinquefasciatus* & WNV"
  )) %>% 
  distinct() %>%
  # Arrange along the plotting variables
  arrange(system_ID, sigmaH, KH, mean) %>% 
  mutate(sigmaH_proxy = sigmaH)

@@

  # fill:
  scale_fill_manual(
    name = "Transmission\nthermal optimum (°C)",
    values = met.brewer("Johnson",
                        n = length(Topt_breaks),
                        direction = -1
    )
  ) +
  # color:
  scale_colour_manual(
    name = "",
    labels = c("R0=1" = unname(TeX("$\\mathcal{R}_0=1$"))),
    values = c("black"),
    guide = guide_legend(
      order = 1,
      keyheight = unit(1, "cm"),
      override.aes = list(size = 6)
    )
  ) +
  # faceting:
  facet_wrap(~system_label, scales = "free", nrow = 2, labeller = labeller()) +
  # legend:
  guides(fill = guide_coloursteps(
    title.position = "top", title.hjust = 0,
    barheight = unit(5, "cm"),
    show.limits = TRUE
  )) +
  # theme options:
  theme_minimal(16) +
  theme(
    legend.box = "vertical",
    legend.position = "right",
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.direction = "horizontal",
    legend.justification = "left",
    axis.title.x = element_text(hjust = 0.5),
    strip.text = ggtext::element_markdown(
      size = 11,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )
